PANCHANGE_data_preprocessing

### 2020-08-11 PANCHANGE_data_preprocessing2020-07-15.Rmd
added if statements for education and started reading in the NBR data

# Specify cohort
```{r}
GLAD = TRUE
EDGI = FALSE
NBR = FALSE
RAMP = FALSE
```

@@

# GLAD demographics
## Read in data
```{r glad dem cols}
glad.dem.raw <- readRDS(file = paste0(data.raw_path, "/glad/dem_glad.rds"))
dim(glad.dem.raw)
colnames(glad.dem.raw)
```

# Select variables
```{r glad dem cleaning}
exclude_cols_dem <- c("ID",
                  "Sample",
                  "startDate.glad",
                  "Age_uncleaned",
                  "EduYrs")

@@

# Inspect dimensions
dim(glad.dem.raw.id)
# Inspect colnames
colnames(glad.dem.raw.id)
#Differences
dim(glad.dem.raw)[1]-dim(glad.dem.raw.id)[1]
```

# NBR cohort information
```{r nbr cohort information}
if(NBR == TRUE){
  nbr.cohort.raw <- readxl::read_xlsx(
    path = paste0(data.raw_path, "/nbr/coping_participants_export_20200730.xlsx")
    )

  dim(nbr.cohort)
  colnames(nbr.cohort)
  }
```

@@

# Highest education
Proceed along qualifications fields. If field is not blank, take that as highest qual. If field is blank, check next field. If all fields are blank, NA
Assumes University degree > A level > NVQ > GCSE/O Level/CSE
```{r Highest education}
#create numeric version of the highest education variable
glad.dem.raw.id <- glad.dem.raw.id %>%
  mutate(
    highest_education_numeric =
      case_when(
        demographics.college_or_university_degree_numeric == "1" ~ 4,
        demographics.a_levelsas_levels_or_equivalent_numeric == "1" ~ 3,
        demographics.nvq_or_hnd_or_hnc_or_equivalent_numeric == "1" ~ 2,
        demographics.o_levelsgcses_or_equivalent_numeric == "1" ~ 1,
        demographics.cses_or_equivalent_numeric == "1" ~ 1)
    )

@@

glad.dem.raw.id %>%
  freq(highest_education)
```

Check highest education coding
```{r Highest education check}
glad.dem.raw.id %>%
  select(highest_education,
        demographics.college_or_university_degree_numeric,
        demographics.a_levelsas_levels_or_equivalent_numeric,
        demographics.nvq_or_hnd_or_hnc_or_equivalent_numeric,
        demographics.o_levelsgcses_or_equivalent_numeric,
        demographics.cses_or_equivalent_numeric
  )
```

>>>>

# Specify cohort
```{r Specify cohort}
GLAD = FALSE
EDGI = FALSE
NBR = TRUE
RAMP = FALSE
```

@@

# GLAD demographics
## Read in data
```{r glad dem cols}
if(GLAD == TRUE){
glad.dem.raw <- readRDS(file = paste0(data.raw_path, "/glad/dem_glad.rds"))
dim(glad.dem.raw)
colnames(glad.dem.raw)
}
```

# Select variables
```{r glad dem cleaning}
if(GLAD == TRUE){
exclude_cols_dem <- c("ID",
                  "Sample",
                  "startDate.glad",
                  "Age_uncleaned",
                  "EduYrs")

@@

# Inspect dimensions
dim(glad.dem.raw.id)
# Inspect colnames
colnames(glad.dem.raw.id)
#Differences
dim(glad.dem.raw)[1]-dim(glad.dem.raw.id)[1]
}
```

# NBR demographics

## Read in data
```{r nbr dem cols}
if(NBR == TRUE){
  nbr.dem.raw <- readRDS(file = paste0(data.raw_path, "/nbr/dem_nbr.rds"))
  dim(nbr.dem.raw)
  colnames(nbr.dem.raw)
}
```

# Select variables
```{r nbr dem cleaning}
if(NBR == TRUE){
  exclude_cols_dem <- c("ID",
                        "Sample",
                        "startDate.nbr",
                        "Age_uncleaned",
                        "EduYrs")

  nbr.dem.raw.id <- nbr.dem.raw %>%
    drop_na(subjectid) %>% # Drop NAs
    distinct(subjectid, .keep_all = TRUE) %>% # Remove duplicates based on ID
    separate(subjectid, into = c("Sample", "ID.intm"), sep = 6) %>% # Split ID in Sample and Number
    separate(ID.intm, into = "ID", sep = 7) %>%
    mutate(ID = as.numeric(ID),
           startDate.coping = startDate) %>% # Rename startDate
    select(
      Sample, # Sample
      ID, # ID
      startDate.coping, #Date when participant signed up to nbr
#      Age_uncleaned = demographics.how_old_are_you_now.txt, # Age
      Birthyear = demographics.year, # Birthyear
      Gender = demographics.which_gender_do_you_identify_with, # Gender
      Sex = demographics.select_questionnaire_items_medical, #Sex
      EduYrs = demographics.school_education_school_years.txt, # Years in education
      demographics.college_or_university_degree, # level of education - univeristy
      demographics.a_levelsas_levels_or_equivalent, # level of education - alevel
      demographics.o_levelsgcses_or_equivalent, # level of education - gcse
      demographics.cses_or_equivalent, # level of education - cse
      demographics.nvq_or_hnd_or_hnc_or_equivalent, # level of education - nvq
      Ethnicity = demographics.questions_based_ethnic_origin # Ethnicity
    ) %>%
    add_numeric(., exclude = exclude_cols_dem) %>%
    mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
    mutate_if(is.numeric, ~na_if(., -99)) %>%
    mutate_if(is.numeric, ~na_if(., -77)) %>%
    mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
    mutate_if(is.factor, ~na_if(., "Don't know")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
    mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
    mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

  # Inspect dimensions
  dim(nbr.dem.raw.id)
  # Inspect colnames
  colnames(nbr.dem.raw.id)
  #Differences
  dim(nbr.dem.raw)[1]-dim(nbr.dem.raw.id)[1]
}
```

## Cohort information
Information on INTERVAL, COMPARE, and STRIDES
http://www.donorhealth-btru.nihr.ac.uk/studies/
```{r nbr cohort information}
if(NBR == TRUE){
  nbr.cohort.raw <- readxl::read_xlsx(
    path = paste0(data.raw_path, "/nbr/coping_participants_export_20200730.xlsx")
    )

  dim(nbr.cohort.raw)
  colnames(nbr.cohort.raw)
  }
```

@@

# Highest education
Proceed along qualifications fields. If field is not blank, take that as highest qual. If field is blank, check next field. If all fields are blank, NA
Assumes University degree > A level > NVQ > GCSE/O Level/CSE
```{r Highest education}
if(GLAD == TRUE | EDGI == TRUE) {
#create numeric version of the highest education variable
glad.dem.raw.id <- glad.dem.raw.id %>%
  mutate(
    highest_education_numeric =
      case_when(
        demographics.college_or_university_degree_numeric == "1" ~ 4,
        demographics.a_levelsas_levels_or_equivalent_numeric == "1" ~ 3,
        demographics.nvq_or_hnd_or_hnc_or_equivalent_numeric == "1" ~ 2,
        demographics.o_levelsgcses_or_equivalent_numeric == "1" ~ 1,
        demographics.cses_or_equivalent_numeric == "1" ~ 1)
    )

@@

glad.dem.raw.id %>%
  freq(highest_education)
}
```

Check highest education coding
```{r Highest education check}
if(GLAD == TRUE | EDGI == TRUE) {
glad.dem.raw.id %>%
  select(highest_education,
        demographics.college_or_university_degree_numeric,
        demographics.a_levelsas_levels_or_equivalent_numeric,
        demographics.nvq_or_hnd_or_hnc_or_equivalent_numeric,
        demographics.o_levelsgcses_or_equivalent_numeric,
        demographics.cses_or_equivalent_numeric
  )
}
```
